```{r,echo=FALSE,results='hide',warning=FALSE}
# Make sure the working directory is set correctly
setwd("C:/Users/Reese Anschultz/eda-ProgrammingAssignment2")
```
---
title: "Reproducible Research: Peer Assessment 1"
output: 
  html_document:
    keep_md: true
---

If you haven't done so already, install ggplot2 using:
```
install.packages("ggplot2")
```

## Loading and preprocessing the data
```{r,warning=FALSE}
library(ggplot2)

zip_file<-"activity.zip"
activity_data_file<-"activity.csv"
unzip(zip_file,junkpaths=TRUE)
activity_data<-read.csv(activity_data_file,header=TRUE,na.strings="NA")
activity_data$date<-as.Date(activity_data$date)
```

## What is mean total number of steps taken per day?
1. Calculate the total number of steps taken per day
```{r}
activity_data_per_day<-aggregate(activity_data['steps'],activity_data['date'],FUN=sum)
```

2. Make a histogram of the total number of steps taken each day
```{r,warning=FALSE}
ggplot(data=activity_data_per_day)+geom_histogram(aes(x=date,y=steps),stat="identity")
```

3. Calculate and report the mean and median of the total number of steps taken per day  

```{r}
mean_steps_per_day<-sapply(activity_data_per_day['steps'],mean,na.rm=TRUE)
```
The mean number of steps taken each day is `r format(mean_steps_per_day,digits=7,nsmall=2)`.

```{r}
median_steps_per_day<-sapply(activity_data_per_day['steps'],median,na.rm=TRUE)
```
The median number of steps taken each day is `r format(median_steps_per_day,digits=2,nsmall=2)`.


## What is the average daily activity pattern?
1. Make a time series plot (i.e. type = "l") of the 5-minute interval (x-axis) and the average number of steps taken, averaged across all days (y-axis)
```{r}
mean_steps_per_interval<-aggregate(activity_data['steps'],activity_data['interval'],FUN=sum,na.rm=TRUE)
ggplot(data=mean_steps_per_interval)+geom_line(aes(x=interval,y=steps))
```

2. Which 5-minute interval, on average across all the days in the dataset, contains the maximum number of steps?

The interval with the maximum number of steps, on average, is `r mean_steps_per_interval[which.max(mean_steps_per_interval$steps),"interval"]`.

## Imputing missing values



## Are there differences in activity patterns between weekdays and weekends?
